---
title: "Project"
author: "Angad Singh"
date: "17/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Business Questions
```{r}
# Business questions we can ask:

# Are cancellation patterns different across different market segments?

# Are cancellation patterns different across different deposit types? 

# Are there additional kinds of data that could be useful for the hotel to 
# collect and how? 

# Why might guests who have previously stayed at the hotel cancel their 
# bookings? 

# Are cancellations different across the top three most common countries? 
```

Creating Our Datasets
```{r}
library(tidyverse)

bookings_full <-  read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/Resort01.csv")

bookings_cleaned <- bookings_full %>% filter(Country != "NULL")

bookings_cleaned$TotalNights <- bookings_cleaned$StaysInWeekNights + bookings_cleaned$StaysInWeekendNights

bookings_canceled <- bookings_cleaned %>% filter(IsCanceled == 1)

bookings_stayed <- bookings_cleaned %>% filter(IsCanceled == 0)
```

Checking for Missing Values
```{r}
sapply(bookings_full, function(x) sum(is.na(x)))
```

Examining the Structure of Our Datasets
```{r}
str(bookings_cleaned)

summary(bookings_cleaned)

# The following line was used per column to confirm the column's data type
mode(bookings_cleaned$Meal)
```

```{r}
# There are 11,077 canceled bookings and 28519 bookings. 
```

Data Exploration
```{r}
table(bookings_cleaned$ReservedRoomType)

rooms_list <- unique(bookings_cleaned[c("ReservedRoomType")])
```

```{r}
table(bookings_cleaned$AssignedRoomType)

rooms_list <- unique(bookings_cleaned[c("AssignedRoomType")])
```

```{r}
table(bookings_canceled$Meal)
```

```{r}
table(bookings_cleaned$Meal)
```

```{r}
summary(bookings_canceled$LeadTime)

range(bookings_canceled$LeadTime)
```

```{r}
table(bookings_cleaned$DepositType)

table(bookings_canceled$DepositType)
```

```{r}
summary(bookings_canceled$BookingChanges)

range(bookings_canceled$BookingChanges)
```

```{r}
range(bookings_cleaned$PreviousCancellations)

range(bookings_cleaned$PreviousBookingsNotCanceled)

range(bookings_cleaned$BookingChanges)

table(bookings_cleaned$CustomerType)

range(bookings_cleaned$RequiredCarParkingSpaces)
```

Cancelled Bookings by Lead Time
```{r}
same_day_cancelled <- bookings_canceled %>% filter(LeadTime == 0)
```

Bookings by Previous Guest 
```{r}
guest_previous <- bookings_cleaned %>% filter(IsRepeatedGuest == 1)

guest_new <- bookings_cleaned %>% filter(IsRepeatedGuest == 0)

guest_previous[guest_previous$IsCanceled == 1, ]

# There are 111 bookings for repeated guests who cancelled their stays

range(bookings_cleaned$IsRepeatedGuest)

bookings_cleaned %>% filter(IsRepeatedGuest == 1)

sum(bookings_cleaned$IsRepeatedGuest == 1)

sum(bookings_canceled$IsRepeatedGuest == 1) / sum(bookings_cleaned$IsRepeatedGuest == 1) 

max(bookings_cleaned$IsRepeatedGuest)

summary(bookings_cleaned$PreviousCancellations)

bookings_cleaned[bookings_cleaned$PreviousCancellations == 26, c("PreviousCancellations")]

max(bookings_cleaned$PreviousCancellations)

summary(bookings_cleaned$PreviousBookingsNotCanceled)

bookings_cleaned[bookings_cleaned$PreviousBookingsNotCanceled == 30, ]

max(bookings_cleaned$PreviousBookingsNotCanceled)

range(bookings_cleaned$BookingChanges)
```

Exploring the Country Column
```{r}
countries_list <- unique(bookings_full[c("Country")])

country_null <- bookings_full %>% filter(Country == "NULL")
```

```{r}
bookings_PRT <- bookings_cleaned %>% filter(Country == "PRT")

bookings_not_PRT <- bookings_cleaned %>% filter(Country != "PRT")
```

```{r}
(sum(bookings_PRT$IsCanceled == 1) / 17630) * 100

(sum(bookings_not_PRT$IsCanceled == 1) / 21966) * 100
```

```{r}
table(bookings_cleaned$Country)
```
```{r}
table(bookings_canceled$Country)
```

```{r}
country_all <- data.frame(table(bookings_cleaned$Country))

country_cancel <- data.frame(table(bookings_canceled$Country))
```

Exploring Market Segments
```{r}
# What are the different market segments? 

market_segments <- unique(bookings_cleaned[c("MarketSegment")])
market_segments
```
```{r}
# There are 6 different kinds of market segments: Direct, Corporate, Online TA, 
# Offline TA/TO, Complementary, and Groups 

bookings_cleaned[bookings_cleaned$MarketSegment == "Direct", ]
# There are 6,357 total "Direct" bookings

bookings_cleaned[bookings_cleaned$MarketSegment == "Corporate", ]
# There are 2,126 total "Corporate" bookings

bookings_cleaned[bookings_cleaned$MarketSegment == "Online TA", ]
# There are 17,655 total "Online TA" bookings

bookings_cleaned[bookings_cleaned$MarketSegment == "Offline TA/TO", ]
# There are 7,427 total "Offline TA/TO" bookings

bookings_cleaned[bookings_cleaned$MarketSegment == "Complementary", ]
# There are 199 total "Complementary" bookings

bookings_cleaned[bookings_cleaned$MarketSegment == "Groups", ]
# There are 5,832 total "Groups" bookings
```

```{r}
bookings_canceled %>% filter(MarketSegment == "Complementary")

bookings_canceled %>% filter(MarketSegment == "Corporate")

bookings_canceled %>% filter(MarketSegment == "Direct")

bookings_canceled %>% filter(MarketSegment == "Groups")

bookings_canceled %>% filter(MarketSegment == "Offline TA/TO")

bookings_canceled %>% filter(MarketSegment == "Online TA")
```

Market Segment Graphs
```{r}
library(ggplot2)
library(dplyr)

market_canc_tot <- bookings_cleaned %>% group_by(MarketSegment) %>% summarize(total = sum(IsCanceled))

mode(market_canc_tot)

bar_market <- ggplot(market_canc_tot, aes(x = MarketSegment, y = total)) + geom_col(color = "black", fill = "red")
  
bar_market <- bar_market + ggtitle("Total Number of Cancellations per Market Segment") + 
    ylab("Number of Cancellations") +
    xlab("Market Segment")

bar_market
```

```{r}
perc_direct <- sum(bookings_canceled$MarketSegment == "Direct") / sum(bookings_cleaned$MarketSegment == "Direct") * 100

perc_corp <- sum(bookings_canceled$MarketSegment == "Corporate") / sum(bookings_cleaned$MarketSegment == "Corporate") * 100

perc_online_TA <- sum(bookings_canceled$MarketSegment == "Online TA") / sum(bookings_cleaned$MarketSegment == "Online TA") * 100

perc_offline_TA <- sum(bookings_canceled$MarketSegment == "Offline TA/TO") / sum(bookings_cleaned$MarketSegment == "Offline TA/TO") * 100

perc_comp <- sum(bookings_canceled$MarketSegment == "Complementary") / sum(bookings_cleaned$MarketSegment == "Complementary") * 100

perc_groups <- sum(bookings_canceled$MarketSegment == "Groups") / sum(bookings_cleaned$MarketSegment == "Groups") * 100
```

```{r}
perc_direct_canc <- (sum(bookings_canceled$MarketSegment == "Direct") / 11077) * 100

perc_corp_canc <- (sum(bookings_canceled$MarketSegment == "Corporate") / 11077) * 100

perc_online_TA_canc <- (sum(bookings_canceled$MarketSegment == "Online TA") / 11077) * 100

perc_offline_TA_canc <- (sum(bookings_canceled$MarketSegment == "Offline TA/TO") / 11077) * 100

perc_comp_canc <- (sum(bookings_canceled$MarketSegment == "Complementary") / 11077) * 100

perc_groups_canc <- (sum(bookings_canceled$MarketSegment == "Groups") / 11077) * 100
```

```{r}
cancelled_tot_by_segment <- data.frame(MarketSegment = unique(bookings_cleaned[c("MarketSegment")]))

cancelled_tot_by_segment$PercentCanceled <- c(perc_direct_canc, perc_corp_canc, perc_online_TA_canc, perc_offline_TA_canc, perc_comp_canc, perc_groups_canc)
```

How many cancellations are there per market segment?
```{r}
bar_market_canc_perc <- ggplot(cancelled_tot_by_segment, 
                          aes(x = MarketSegment, y = PercentCanceled)) + 
  geom_col(color = "black", fill = "red")
  
bar_market_canc_perc <- bar_market_canc_perc + ggtitle("% of Total Cancellations per Market Segment") + 
    ylab("% of Total Cancellations") +
    xlab("Market Segment")

bar_market_canc_perc
```

```{r}
cancelled_by_segment <- data.frame(MarketSegment = unique(bookings_cleaned[c("MarketSegment")]))

cancelled_by_segment$PercentCanceled <- c(perc_direct, perc_corp, perc_online_TA, perc_offline_TA, perc_comp, perc_groups)

view(cancelled_by_segment)
```

What market segments contribute the most to cancellations overall?
```{r}
bar_market_perc <- ggplot(cancelled_by_segment, 
                          aes(x = MarketSegment, y = PercentCanceled)) + 
  geom_col(color = "black", fill = "red")
  
bar_market_perc <- bar_market_perc + ggtitle("% of Cancellations per Total Market Segment Bookings") + 
    ylab("% of Cancellations") +
    xlab("Market Segment")

bar_market_perc
```

Exploring Number of Adult Guests
```{r}
range(bookings_cleaned$Adults)

sum(bookings_canceled$Adults == 1) / sum(bookings_cleaned$Adults == 1) * 100
```

```{r}
mean(bookings_cleaned$Adults)

bookings_cleaned[bookings_cleaned$MarketSegment == "Direct", ]

bookings_cleaned[bookings_cleaned$Adults == 2, ]
# There are 31,258 bookings with 2 adult guests

(31258 / 39596) * 100
# ~79% of all bookings have 2 adult guests

bookings_adults_only <- bookings_cleaned %>% filter(Adults == 2) %>% filter(Children == 0) %>% filter(Babies == 0)
# There are 27,881 bookings that have only adult guests

(27881 / 39596) * 100
# ~70% of all bookings have exactly 2 adult guests 

(sum(bookings_adults_only$IsCanceled == 1) / 27881) * 100 

range(bookings_adults_only$Children)

range(bookings_adults_only$Babies)

bookings_adults_only_canc <- bookings_canceled %>% filter(Adults == 2) %>% filter(Children == 0) %>% filter(Babies == 0)
# There are 8306 canceled bookings with exactly 2 adult guests

(8306 / 11077) * 100
# ~75% of all cancellations have exactly 2 adult guests 
```

```{r}
perc_adults_1 <- sum(bookings_canceled$Adults == 1) / sum(bookings_cleaned$Adults == 1) * 100

perc_adults_2 <- sum(bookings_canceled$Adults == 2) / sum(bookings_cleaned$Adults == 2) * 100

perc_adults_3 <- sum(bookings_canceled$Adults == 3) / sum(bookings_cleaned$Adults == 3) * 100
```

```{r}
cancelled_by_num_adults <- data.frame(c(1, 2, 3))

cancelled_by_num_adults$PercentCanceled <- c(perc_adults_1, perc_adults_2, perc_adults_3)
```

```{r}
bar_adults_perc <- ggplot(cancelled_by_num_adults, 
                          aes(x = c.1..2..3., y = PercentCanceled)) + 
  geom_col(color = "black", fill = "red")
  
bar_adults_perc <- bar_adults_perc + ggtitle("% of Cancellations per Total Bookings by Number of Adult Guests") + 
    ylab("% of Cancellations") +
    xlab("Number of Adult Guests")

bar_adults_perc
```

```{r}
counts_adults <- table(bookings_cleaned$IsCanceled, bookings_cleaned$Adults)

bar_adults <- barplot(counts_adults, main = "Bookings by Number of Adult Guests", col = c("darkblue", "red"),
        legend = c("Not Cancelled", "Cancelled"), xlab = ("# of Adult Guests"), ylab = ("# of Bookings"))
```

```{r}
bookings_cleaned %>% filter(Adults == 1 | Adults == 2 | Adults == 3)

bookings_cleaned %>% filter(Adults == 1) 

bookings_cleaned %>% filter(Adults == 2)

bookings_cleaned %>% filter(Adults == 3)

39538 / 39596 * 100
```

```{r}
perc_adults_1_canc <- (sum(bookings_canceled$Adults == 1) / 11077) * 100

perc_adults_2_canc <- (sum(bookings_canceled$Adults == 2) / 11077) * 100

perc_adults_3_canc <- (sum(bookings_canceled$Adults == 3) / 11077) * 100
```

```{r}
cancelled_by_canc_adults <- data.frame(c(1, 2, 3))

cancelled_by_canc_adults$PercentCanceled <- c(perc_adults_1_canc, perc_adults_2_canc, perc_adults_3_canc)
```

```{r}
bar_adults_perc_canc <- ggplot(cancelled_by_canc_adults, 
                          aes(x = c.1..2..3., y = PercentCanceled)) + 
  geom_col(color = "black", fill = "red")
  
bar_adults_perc_canc <- bar_adults_perc_canc + ggtitle("% of Cancellations per Cancelled Bookings by Number of Adult Guests") + 
    ylab("% of Cancellations") +
    xlab("Number of Adult Guests")

bar_adults_perc_canc
```

Creating Specific Dataframes for Market Segments

Online TA
```{r}
bookings_online_TA <- bookings_cleaned %>% filter(MarketSegment == "Online TA")
```

```{r}
summary(bookings_online_TA$Adults)

range(bookings_online_TA$Adults)

summary(bookings_online_TA$Children)

range(bookings_online_TA$Children)

summary(bookings_online_TA$Babies)

range(bookings_online_TA$Babies)

summary(bookings_online_TA$StaysInWeekNights)

range(bookings_online_TA$StaysInWeekNights)
```

Groups
```{r}
bookings_groups <- bookings_cleaned %>% filter(MarketSegment == "Groups")
```

```{r}
summary(bookings_groups$Adults)

range(bookings_groups$Adults)

summary(bookings_groups$Children)

range(bookings_groups$Children)

summary(bookings_groups$Babies)

range(bookings_groups$Babies)

summary(bookings_groups$StaysInWeekNights)

range(bookings_groups$StaysInWeekNights)
```

Creating Box Plots for Different Market Segments
```{r}
boxplot(bookings_groups$StaysInWeekNights)

boxplot(bookings_online_TA$StaysInWeekNights)
```

Creating a Bar Plot for Group Cancellations by Deposit Type
```{r}
bookings_groups_canc <- bookings_canceled %>% filter(MarketSegment == "Groups") 

table(bookings_groups_canc$Country)

table(bookings_groups_canc$DepositType)

groups_canc_deposit <- data.frame(c(1007, 1454, 13))

groups_canc_deposit$Deposit <- (c("No Deposit", "Non-Refundable", "Refundable")) 
```

```{r}
bar_groups_deposit_canc <- ggplot(groups_canc_deposit, 
                          aes(x = Deposit , y = c.1007..1454..13.)) + 
  geom_col(color = "black", fill = "red")
  
bar_groups_deposit_canc <- bar_groups_deposit_canc + ggtitle("Number of Groups Cancellations per Deposit Type") + 
    ylab("Number of Cancellations") +
    xlab("Number of Adult Guests")

bar_groups_deposit_canc
```

Examining Cancellations for Online TA Bookings
```{r}
bookings_online_canc <- bookings_canceled %>% filter(MarketSegment == "Online TA") 

table(bookings_online_canc$Country)

table(bookings_online_canc$DepositType)
```


```{r}

library(tidyverse)
library(tidyverse)

bookings_full <-  read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/Resort01.csv")

bookings_cleaned <- bookings_full %>% filter(Country != "NULL")

bookings_cleaned$TotalNights <- bookings_cleaned$StaysInWeekNights + bookings_cleaned$StaysInWeekendNights

bookings_canceled <- bookings_cleaned %>% filter(IsCanceled == 1)

bookings_stayed <- bookings_cleaned %>% filter(IsCanceled == 0)

counts_tot_nights <- table(bookings_cleaned$IsCanceled, 
                            bookings_cleaned$StaysInWeekendNights + bookings_cleaned$StaysInWeekNights)

# view(bookings_cleaned)

bar_total_nights <- barplot(counts_tot_nights, main = "Total Nights", 
            col = c("darkblue", "red"),legend = c("Not Cancelled", "Cancelled"))

mean(bookings_cleaned$TotalNights)

mean(bookings_canceled$TotalNights)

datacancelled <- bookings_cleaned[bookings_cleaned$IsCanceled == 1,]
datanocancel <- bookings_cleaned[bookings_cleaned$IsCanceled == 0,]

```

## Plots




LeadTime:
```{r}
countsLeadTime <- table(bookings_cleaned$IsCanceled, bookings_cleaned$LeadTime)
```

Barplot
```{r}
barLeadTime <- barplot(countsLeadTime, main ="Bookings by LeadTime", 
            col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
            xlab = ("LeadTime"), las=1, beside = TRUE)
boxLeadtime <- ggplot(bookings_cleaned, aes(x = as.factor(IsCanceled), 
                                    y = LeadTime)) + 
              geom_boxplot() + coord_flip()
```


Weekend Nights, Week Nights, Total Nights
```{r}
countsWeekendNights <- table(bookings_cleaned$IsCanceled, bookings_cleaned$StaysInWeekendNights)

countsWeekNights <- table(bookings_cleaned$IsCanceled, bookings_cleaned$StaysInWeekNights)

countsTotalNights <- table (bookings_cleaned$IsCanceled, 
                            bookings_cleaned$StaysInWeekendNights+bookings_cleaned$StaysInWeekNights)
```

Barplot
```{r}
barWeekend <- barplot(countsWeekendNights, main = "Bookings by Number Weekend Nights",
            col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
            xlab = ("# of Weekend Nights"), las=1, beside = TRUE)



barWeek <- barplot(countsWeekNights, main = "Bookings by Weeknights", 
            col = c("darkblue", "red"),legend = c("Not Cancelled", "Cancelled"),
            xlab = ("# of Weeknights"), las=1, beside = TRUE)



barTotal <- barplot(countsTotalNights, main = "Bookings by Total Nights", 
            col = c("darkblue", "red"),legend = c("Not Cancelled", "Cancelled"),
            xlab = ("# of Total Nights"),
            las=1, beside = TRUE)
```


Adults, Children, Babies
```{r}
countsAdults <- table(bookings_cleaned$IsCanceled, bookings_cleaned$Adults)

countsChildren <- table(bookings_cleaned$IsCanceled, bookings_cleaned$Children)

countsBabies <- table(bookings_cleaned$IsCanceled, bookings_cleaned$Babies)
```

Barplot
```{r}
barAdults <- barplot(countsAdults, main = "Bookings by Number of Adult Guests", col = c("darkblue", "red"),
        legend = c("Not Cancelled", "Cancelled"), xlab = ("# of Adult Guests"), las=1, beside = TRUE)

barBabies <- barplot(countsBabies, main = "Bookings by 
                     Number of Babies", col = c("darkblue", "red"),
        legend = c("Not Cancelled", "Cancelled"), xlab = ("# of Babies"), 
        las=1, beside = TRUE)


barChildren <- barplot(countsChildren, main = "Bookings by Number of Children",
            col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
            xlab = ("# of Children"),
            las=1, beside = TRUE)
```

Meal
```{r}
countsMeals <- table(bookings_cleaned$IsCanceled, bookings_cleaned$Meal)
```
Barplot
```{r}
barMeals <- barplot(countsMeals, main = "Bookings by Meal Type",
            col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
            xlab = ("Meal Type"),
            las=1, beside = TRUE)
```


Country
```{r}
countsCountry <- table(bookings_cleaned$IsCanceled, bookings_cleaned$Country)
```

Cancel Ratio:
```{r}

```



Add Map Please:
```{r}

```


Repeated Guest
```{r}

countsRepeatedGuest <- table(bookings_cleaned$IsCanceled,
                             bookings_cleaned$IsRepeatedGuest)
```

Barplot
```{r}
barRepeated <- barplot(countsRepeatedGuest, main = "Bookings by Repeated Customer", 
             col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
             xlab = ("Repeated Customer"), las=1, beside = TRUE)
```


Previous Cancellations / Not Cancellation
```{r}
countsPrevCancel <- table(bookings_cleaned$IsCanceled,bookings_cleaned$PreviousCancellations)
bookings_cleaned$History_Cancel <- bookings_cleaned$PreviousBookingsNotCanceled - bookings_cleaned$PreviousCancellations < 0
countsHistory_Cancel <- table(bookings_cleaned$IsCanceled, bookings_cleaned$History_Cancel)
```

Barplot
```{r}
barPrevCancel <- barplot(countsPrevCancel, main = "Bookings by Previous Cancellations", 
            col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
            xlab = ("Previous Cancellations"), las=1, beside = TRUE)

bookings_cleaned$History_Cancel = as.factor(bookings_cleaned$PreviousBookingsNotCanceled - bookings_cleaned$PreviousCancellations < 0)

barHistoryCancel <- barplot(countsHistory_Cancel, 
            main = "Bookings by History of Cancellation", 
            col = c("darkblue", "red"),
        legend = c("Not Cancelled", "Cancelled"), 
        xlab = ("History of Cancellation"),
        las=1, beside = TRUE)
```

Reserved Room, Assigned Room
```{r}
countsReservedRoom <- table(bookings_cleaned$IsCanceled, bookings_cleaned$ReservedRoomType)
countsAssignedRoom <- table(bookings_cleaned$IsCanceled, bookings_cleaned$AssignedRoomType)
bookings_cleaned$Get_Desired_Room = bookings_cleaned$ReservedRoomType == bookings_cleaned$AssignedRoomType
countsGetDesiredRoom <- table(bookings_cleaned$IsCanceled, bookings_cleaned$Get_Desired_Room)

```
Barplots:
```{r}
barReservedRoom <- barplot(countsReservedRoom, 
            main = "Bookings by Reserved Room Type", 
            col = c("darkblue", "red"),
        legend = c("Not Cancelled", "Cancelled"), 
        xlab = ("Reserved Room Type"),
        las=1, beside = TRUE)

barAssignedRoom <- barplot(countsAssignedRoom, 
            main = "Bookings by Assigned Room Type", 
            col = c("darkblue", "red"),
        legend = c("Not Cancelled", "Cancelled"), 
        xlab = ("Assigned Room Type"),
        las=1, beside = TRUE)

barGetDesiredRoom <- barplot(countsGetDesiredRoom, 
            main = "Bookings by Whether or not Client got their Reserved Room", 
            col = c("darkblue", "red"),
        legend = c("Not Cancelled", "Cancelled"), 
        xlab = ("Whether or not Client got their Reserved Room"),
        las=1, beside = TRUE)
```



Booking Changes
```{r}
countsBookingChanges <- table(bookings_cleaned$IsCanceled, bookings_cleaned$BookingChanges)
```
Barplot:
```{r}
barBookingChanges <- barplot(countsBookingChanges, 
            main = "Reservations by Number of Booking Changes", 
            col = c("darkblue", "red"),
        legend = c("Not Cancelled", "Cancelled"), 
        xlab = ("Number of Booking Changes"),
        las=1, beside = TRUE)
```



Parking Requirements
```{r}
countsReqParking <- table(bookings_cleaned$IsCanceled, bookings_cleaned$RequiredCarParkingSpaces)
```

Barplot
```{r}
barParking <- barplot(countsReqParking, main = "Bookings by Parking Requests", 
             col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
             xlab = ("# of Parking Requests"),
             las=1, beside = TRUE)
```


Special Requests
```{r}
countsSpecReq <- table(bookings_cleaned$IsCanceled, bookings_cleaned$TotalOfSpecialRequests)
```

Barplot
```{r}
barSpecReq <- barplot(countsSpecReq, main = "Bookings by Special Requirements", 
             col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
             xlab = ("# of Special Requirements"),
             las=1, beside = TRUE)
```


Deposit (Yes/No; Refundable/Not)
```{r}
countsDeposit <- table(bookings_cleaned$IsCanceled, as.factor(bookings_cleaned$DepositType))
```

Barplot
```{r}
barDeposit <- barplot(countsDeposit, main = "Bookings by Deposit Type", 
             col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
             xlab = ("Deposit Type"),
             las=1, beside = TRUE)
```

Market Segment
```{r}
countsMarketSeg <- table(bookings_cleaned$IsCanceled, as.factor(bookings_cleaned$MarketSegment))
```

Barplot
```{r}
barMarkSeg <- barplot(countsMarketSeg, main = "Bookings by Market Segment", 
              col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
              xlab = ("Market Segment"),
              las=1, beside = TRUE)
```


Customer Type
```{r}
countsCustomerType <- table(bookings_cleaned$IsCanceled, as.factor(bookings_cleaned$CustomerType))
```

Barplot
```{r}
barCustType <- barplot(countsCustomerType, main = "Bookings by Customer Type", 
      col = c("darkblue", "red"), legend = c("Not Cancelled", "Cancelled"),
      xlab = ("Customer Type"),
      las=1, beside = TRUE)
```





Rules:

Data conversion for generating Rules:
```{r}
library(arules)
library(arulesViz)

data_new <- data.frame(Cancellation = as.factor(bookings_cleaned$IsCanceled),
                       LeadTime = as.factor(bookings_cleaned$LeadTime),
                       WeekendNights = as.factor(bookings_cleaned$StaysInWeekendNights),
                       WeekNights = as.factor(bookings_cleaned$StaysInWeekNights),
 TotalNights = as.factor(bookings_cleaned$StaysInWeekendNights+bookings_cleaned$StaysInWeekNights),
                        Adults = as.factor(bookings_cleaned$Adults),
 Babies = as.factor(bookings_cleaned$Babies),
 Chidren = as.factor(bookings_cleaned$Children), Meal = as.factor(bookings_cleaned$Meal),
 Country = as.factor(bookings_cleaned$Country), Market_Seg = as.factor(bookings_cleaned$MarketSegment),
 Repeated_Guest = as.factor(bookings_cleaned$IsRepeatedGuest), Prev_Cancel = 
   as.factor(bookings_cleaned$PreviousCancellations), PrevBook = as.factor(bookings_cleaned$PreviousBookingsNotCanceled), History_Cancel = as.factor(bookings_cleaned$PreviousBookingsNotCanceled - bookings_cleaned$PreviousCancellations < 0),
 Reserved_Room = as.factor(bookings_cleaned$ReservedRoomType), 
 Assigned_Room = as.factor(bookings_cleaned$AssignedRoomType),
 Get_Desired_Room = bookings_cleaned$ReservedRoomType == bookings_cleaned$AssignedRoomType,
 Booking_Changes = as.factor(bookings_cleaned$BookingChanges), 
 Deposit = as.factor(bookings_cleaned$DepositType), 
 Customer_Type = as.factor(bookings_cleaned$CustomerType),
 Car_Park_Required = as.factor(bookings_cleaned$RequiredCarParkingSpaces),
 Special_Requests = as.factor(bookings_cleaned$TotalOfSpecialRequests))

str(data_new)
```

Generating rules with a Confidence = 0.85 and Support = 0.05:

```{r}
datax <- as(data_new, Class = "transactions")
# itemFrequency(datax)
# itemFrequencyPlot(datax)
rules <- apriori(datax, parameter = list(supp = 0.05
                                         , conf = 0.85),
                 control =list(verbose = F),
                 appearance = list(default = "lhs", rhs = "Cancellation=1"))
inspect(rules)
```

Creating a decision tree:

```{r}
library(rpart)
library(rpart.plot)
rpartModel <- rpart(IsCanceled ~ ., data = bookings_cleaned)
rpartModel
rpart.plot(rpartModel)
```

Converting to a dataframe that has all numeric variables for more consistency while creating models:

```{r}
data_all_numeric <- data.frame(
  Cancellation = as.numeric(bookings_cleaned$IsCanceled),
  LeadTime = as.numeric(bookings_cleaned$LeadTime),
  WeekendNights = as.numeric(bookings_cleaned$StaysInWeekendNights),
  WeekNights = as.numeric(bookings_cleaned$StaysInWeekNights),
  TotalNights = as.numeric(bookings_cleaned$StaysInWeekendNights+bookings_cleaned$StaysInWeekNights),
  Adults = as.numeric(bookings_cleaned$Adults),
  Babies = as.numeric(bookings_cleaned$Babies),
  Chidren = as.numeric(bookings_cleaned$Children), 
  Meal = as.numeric(as.factor(bookings_cleaned$Meal)),
  Country = as.numeric(as.factor(bookings_cleaned$Country)), 
  Market_Seg = as.numeric(as.factor(bookings_cleaned$MarketSegment)),
  Repeated_Guest = as.numeric(bookings_cleaned$IsRepeatedGuest), 
  Prev_Cancel = as.numeric(bookings_cleaned$PreviousCancellations), 
  PrevBook = as.numeric(bookings_cleaned$PreviousBookingsNotCanceled), 
  History_Cancel = as.numeric(bookings_cleaned$PreviousBookingsNotCanceled - bookings_cleaned$PreviousCancellations < 0),
  Reserved_Room = as.numeric(as.factor(bookings_cleaned$ReservedRoomType)), 
  Assigned_Room = as.numeric(as.factor(bookings_cleaned$AssignedRoomType)),
  Get_Desired_Room = as.numeric(bookings_cleaned$ReservedRoomType == bookings_cleaned$AssignedRoomType),
  Booking_Changes = as.numeric(bookings_cleaned$BookingChanges), 
  Deposit = as.numeric(as.factor(bookings_cleaned$DepositType)), 
  Customer_Type = as.numeric(as.factor(bookings_cleaned$CustomerType)),
  Car_Park_Required = as.numeric(bookings_cleaned$RequiredCarParkingSpaces),
  Special_Requests = as.numeric(bookings_cleaned$TotalOfSpecialRequests)
  )

str(data_all_numeric)
```

Creating linear models:

First iteration:
```{r message=TRUE}
lmOut <- lm(Cancellation ~ ., data = data_all_numeric)
summary(lmOut)
```

Second iteration:
```{r}
lmOutv2 <- lm(Cancellation ~ LeadTime + WeekendNights + Adults + Chidren + Meal + Country + Market_Seg + Repeated_Guest + Prev_Cancel + History_Cancel + Reserved_Room + Get_Desired_Room + Booking_Changes + Deposit + Customer_Type + Car_Park_Required + Special_Requests, data = data_all_numeric)
summary(lmOutv2)
```

Third iteration:
```{r}
lmOutv3 <- lm(Cancellation ~ LeadTime + Adults + Chidren + Meal + Country + Market_Seg + Repeated_Guest + Prev_Cancel + History_Cancel + Get_Desired_Room + Booking_Changes + Deposit + Customer_Type + Car_Park_Required + Special_Requests, data = data_all_numeric)
summary(lmOutv3)
```
```{r}
lmOutv4 <- lm(Cancellation ~ LeadTime + Adults + Chidren + Meal + Country + Market_Seg + Repeated_Guest + Prev_Cancel + History_Cancel + Get_Desired_Room + Booking_Changes + Deposit + Customer_Type + Car_Park_Required + Special_Requests, data = data_all_numeric)
summary(lmOutv4)
```

Attempting Factor Analysis on the dataset:

```{r}
paranOut <- paran::paran(subset(data_all_numeric, select = -c(Cancellation)), iterations=330, graph=TRUE)
paranOut
```

```{r}
prinOut <- psych::principal(subset(data_all_numeric, select = -c(Cancellation)), nfactors = paranOut$Retained, rotate="varimax", scores = TRUE)
prinOut
```

Training and Testing:

```{r}
install.packages("caret")
library(caret)
trainHotelDataList <- createDataPartition(y= data_all_numeric$Cancellation, p=.30, list=FALSE)
str(trainHotelDataList)
```

```{r}
trainHotelDataSet <- data_all_numeric[trainHotelDataList, ]
str(trainHotelDataSet)
```


```{r}
testHotelDataSet <- data_all_numeric[-trainHotelDataList,]
str(testHotelDataSet)
```
Building a KSVM model:

```{r}
install.packages("kernlab")
library(kernlab)
outKSVMv1 <- ksvm(Cancellation ~ ., data = trainHotelDataSet, C=5, cross=3, prob.model = TRUE)
outKSVMv1
```
Making predictions:

```{r}
predOut <- predict(outKSVMv1, newdata = testHotelDataSet, type = 'response')
str(predOut)
```

```{r}
matrixmanual <- table(predOut, testHotelDataSet$Cancellation)
sum(diag(matrixmanual)) / sum(matrixmanual)
```


```{r}
library(caret)
confusionMatrix(predOut, testHotelDataSet$Cancellation)
```

```{r}
predsCompare <- data.frame(c(testHotelDataSet$Cancellation, predOut))
view(predsCompare)
```

